### 2015-03-18 ms.tex
planet properties words

% The rules:
% I guess same as DFM because he'll complain if we don't use the
% character limit.
% New sentences on each line.
% Citation format is \citet{AuthornameYY}.
% References go into papers.bib alphabetically, with ties broken
% chronologically.

%\documentclass[]{aastex}
\documentclass{emulateapj}
%\documentclass[]{emulateapj}

\bibliographystyle{apj}
\usepackage{subfigure}
\usepackage{url}
\usepackage{hyperref}
\usepackage{datetime}
\usepackage{longtable}
%\usepackage{lscape}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{ulem}
\usepackage{bm}
\usepackage[usenames,dvipsnames]{xcolor}

%\usepackage{pdflscape}

% FIGSET-MACROS-BEGIN
\newcommand{\noprint}[1]{}
\newcommand{\figsetstart}{{\bf Fig. Set} }
\newcommand{\figsetend}{}
\newcommand{\figsetgrpstart}{}
\newcommand{\figsetgrpend}{}
\newcommand{\figsetnum}[1]{{\bf #1.}}
\newcommand{\figsettitle}[1]{ {\bf #1} }
\newcommand{\figsetgrpnum}[1]{\noprint{#1}}
\newcommand{\figsetgrptitle}[1]{\noprint{#1}}
\newcommand{\figsetplot}[1]{\noprint{#1}}
\newcommand{\figsetgrpnote}[1]{\noprint{#1}}
% FIGSET-MACROS-END
\usepackage{color}

\newcommand{\ron}{\color{red}}
\newcommand{\bon}{\color{blue}}
\newcommand{\gon}{\color{green}}
\newcommand{\coff}{\color{black}\,}

\newcommand{\rprs}{{$R_p/R_{\star}$}}
\newcommand{\ars}{{$a/R_{\star}$}}

\newcommand{\eg}{{\it e.g.}}
\newcommand{\ie}{{\it i.e.}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\kep}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\Ci}{Campaign~1}
\newcommand{\vsini}{{$V \sin i$}}
\newcommand\teff{\ensuremath{T_\text{eff}}}
\newcommand{\kms}{{km\,s$^{-1}$}}
\newcommand{\gcc}{{g\,cm$^{-3}$}}
\newcommand{\rstar}{{$R_\star$}}
\newcommand{\rhostar}{{$\rho_\star$}}
\newcommand{\mearth}{{M$_\oplus$}}
\newcommand{\rearth}{{R$_\oplus$}}
\newcommand{\rsun}{{R$_\odot$}}
\newcommand{\msun}{{M$_\odot$}}
\newcommand{\mjup}{{M$_\textrm{Jup}$}}
\newcommand{\rjup}{{R$_\textrm{Jup}$}}

\newcommand{\mstar}{{$M_\star$}}
\newcommand{\logg}{{log(g)}}
\newcommand{\mh}{{[M/H]}~}
\newcommand{\feh}{{[Fe/H]}~}
\newcommand{\afe}{{[$\alpha$/Fe]}~}

\newcommand{\dd}{{\ensuremath{\,\mathrm{d}}}}

\newcommand{\paperip}{\citep{Foreman-Mackey15}}
\newcommand{\paperit}{\citet{Foreman-Mackey15}}

\newcommand{\is}{importance sampling}
\newcommand{\Is}{Importance sampling}
%\newcommand{\h2ok2}{{$\rm H_2O-K2$}}

\newcommand{\todo}[3]{{\color{#2} \emph{#1} TO DO: #3}}
\newcommand{\btmtodo}[1]{\todo{BEN}{red}{#1}}
\newcommand{\anytodo}[1]{\todo{ANYONE}{NavyBlue}{#1}}
\newcommand{\dfmtodo}[1]{\todo{DFM}{ForestGreen}{#1}}

\newcommand{\Nfp}{4}
\newcommand{\Nvalidated}{21}
\newcommand{\Nlimbo}{11} %adds to 36?  good.

\begin{document}

\title{Stellar and Planetary Properties of \KT\ Campaign 1
Candidates and Validation of \Nvalidated\ Systems, Including a Planet Receiving
Earth-like Insolation}

\newcommand{\caltech}{1}
\newcommand{\cfa}{2}
\newcommand{\nyu}{4}
\newcommand{\princeton}{3}
\newcommand{\mpia}{5}
\newcommand{\cds}{6}
\newcommand{\jcpa}{7}
\newcommand{\texas}{8}
\newcommand{\hjs}{9}

\author{%
% The Heavy Lifters
Benjamin~T.~Montet\altaffilmark{\caltech,\cfa},
Timothy~D.~Morton\altaffilmark{\princeton},
Daniel~Foreman-Mackey\altaffilmark{\nyu},
% The Benefactors
John~Asher~Johnson\altaffilmark{\cfa},
David~W.~Hogg\altaffilmark{\nyu,\mpia,\cds},
% The Observers
% (ranked by integration time x collecting area)
Brendan~P.~Bowler\altaffilmark{\caltech, \jcpa},
David~W.~Latham\altaffilmark{\cfa},
Allyson~Bieryla\altaffilmark{\cfa},
Andrew~W.~Mann\altaffilmark{\texas, \hjs},
And Any TRES Team Members We Should Add
}

\email{btm@astro.caltech.edu}
\altaffiltext{\caltech}  {Cahill Center for Astronomy and Astrophysics,
                          California Institute of Technology, Pasadena, CA,
                          91125, USA}
\altaffiltext{\cfa}      {Harvard-Smithsonian Center for Astrophysics,
                          Cambridge, MA 02138, USA}
\altaffiltext{\princeton}{Department of Astrophysics, Princeton University,
                          Princeton, NJ, 08544, USA}
\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
                          Department of Physics, New York University,
                          4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
                          K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\cds}      {Center for Data Science,
                          New York University,
                          726 Broadway, 7th Floor, New York, NY, 10003, USA}
\altaffiltext{\jcpa} {Caltech Joint Center for Planetary Astronomy Fellow}
\altaffiltext{\texas} {Department of Astronomy, The University of Texas at
                            Austin, Austin, TX 78712, USA}
\altaffiltext{\hjs} {Harlan J. Smith Fellow, The University of Texas at Austin}

\date{\today, \currenttime}

\begin{abstract}
Recently, \paperit{} presented a list of 36 planet candidates orbiting
31 stars in \KT\ \Ci.
In this paper, we present stellar and planetary properties for all systems.
We combine ground-based seeing limited survey data and
adaptive optics imaging with a detailed transit analysis scheme to confirm
\Nvalidated\ as planets and identify \Nfp\ as likely false positives.
Of particular interest is EPIC 201912552, a bright (K=8.9) M2 dwarf
hosting a $2.29 \pm 0.24 R_\odot$ planet with $T_{eq} = 273 \pm 15$ K and an
orbital period of 33 days.
We also present \texttt{vespa}, a new general-purpose implementation
of the \citet{Morton12} procedure to calculate false positive probabilities
for \KT\ candidates.
\end{abstract}

\keywords{catalogs --- planetary systems --- planets and satellites: detection
                   --- stars: fundamental parameters}

\maketitle

\section{Introduction}
The \kep\ telescope \citep{Borucki10} has led to a revolution in stellar and planetary
astrophysics, with 7305 ``objects of interest'' and 4173 ``planet candidates''
discovered to date \citep{Borucki11a, Borucki11b, Batalha13, Burke14, Rowe15, Mullaly15}.
The fidelity of this sample is high, so most of these planet candidates are truly planets
\citep{Morton11b, Fressin13}.
The demise of two reaction wheels on the spacecraft led to a repurposing
of the spacecraft into the \KT\ mission, in which the telescope points at
fields near the ecliptic plane for $\sim 75$ days at a time \citep{Howell14}.
In this observing strategy, two axes of motion of the spacecraft are
controlled by the two remaining reaction wheels, while the roll of the
spacecraft is balanced with solar radiation pressure and quasiperiodic
thruster firing.
The result of this observing strategy is that the targets appear to drift
at the rate of $\sim 1\arcsec$ hr$^{-1}$, with rapid corrections that occur
approximately once every six hours.
Over the full duration of each campaign, the targets appear to remain near the
same location on the detector but both the slow drift and the corrections are
observable by eye \citep{Barentsen15}.

\KT\ light curves produced with aperture photometry contain substantial pointing-induced
photometric variations caused by the star's apparent motion over a poorly-defined flat
field.
Worse yet, these variations occur on timescales similar to transit signals, potentially
masking the observational signature of a planet passing between \kep\ and the host star.

There has been considerable effort to recover these signals.
\citet{Vanderburg14} extract aperture photometry and de-correlate with image centroid
position, ``correcting'' the data to ``remove'' the effects of the spacecraft motion.
These data were used to confirm the first planet discovered in \KT\ \citep{Vanderburg15}.
A similar technique was employed by \citet{Armstrong15a} to identify potential eclipsing
binary systems and to identify a transiting system very near a 3:2 period commensurability
\citep{Armstrong15b}.
\citet{Aigrain15} and \citet{Crossfield15} use a Gaussian Process model for the measured
flux, with pointing measurements as the inputs, and then ``de-trend'' the data with that model.
The latter of these detected the first multi-planet system in \KT\ data.

What is common to all of these methods are that removal of systematics is considered a
step to be undertaken before the search for planets.
Under this strategy, it is implicitly assumed that the systematics are removed perfectly,
while retaining all of the astrophysical signal.
Of course, it is impossible to perfectly separate the astrophysical and instrumental
signal, and such a technique is prone to either over-fitting, in which some of the
astrophysical signal is also removed, or under-fitting, in which some instrumental noise
remains.
A better strategy is to simultaneously fit both the signal and the noise, as is commonly
done in cosmology and, increasingly, in radial velocity searches for planetary systems
\citep[e.g.][]{Ferreira00, Boisse11, Haywood14, Grunblatt15}.

\paperit{} simultaneously fit both the systematics and potential planetary transit signals
in a search for transiting planets.
The authors of that paper assume that the dominant trends in the observed stellar light
curves are caused by spacecraft motion and are shared by many stars.
They then run PCA on all stars to measure the dominant modes, modeling each star as a
linear combination of 150 of these ``eigen light curves'' and a transit signal.
This method enables fitting without over-fitting, and also permits marginalization over
uncertainties induced by the noise model.
Therefore, any uncertainties in the systematics can be propogated into uncertainties in
detected planet parameters, instead of assuming the systematics are understood perfectly.
Using this technique, \paperit{} detect 36 planet candidates orbiting 31 stars in \KT\
\Ci\ data.

In \paperit, only transit properties are provided, not absolute parameters about the
planet or the star.
Additionally, the authors follow the convention of the \kep\ team to include any
transit event as a candidate system rather than a false positive if a secondary eclipse
is not detected: there is no enforced upper limit on the allowed planet radius.
The authors intentionally make no effort to separate astrophysical false positives from
true transiting planets.

In this paper, we present stellar and planetary parameters for each system.
Given these parameters and the transit data, as well as archival imaging, ground-based
seeing limited survey data, and adaptive optics imaging, we are able to confirm
\Nvalidated\ of these systems as transiting planets at the 99\% confidence
level. We also are able to confirm \Nfp\ of these systems as false positives.

This paper is organized as follows.
In \textsection2, we develop stellar properties through
photometric and spectroscopic data.
In \textsection3, we combine the derived stellar properties with \KT\ data to infer planet
candidate properties.
In \textsection4, we combine adaptive optics and radial velocity observations with both
archival and modern ground-based, seeing limited survey data and an analysis of the
transit parameters to calculate false positive probabilities.
In \textsection5, we discuss potentially interesting systems.
In \textsection6, we summarize and discuss our results.

\section{Stellar Properties}
\subsection{Photometry}

With the exception of one star in our sample (EPIC 201912552), we do not have
spectroscopic data with which to characterize the stellar properties.
Instead, we rely on photometry.
For each system, we query the VizieR database of astronomical catalogues
\citep{Ochseinbein00}.
We record the $B$, $V$, $g'$, $r'$, and $i'$ magnitudes and their
uncertainties from the AAVSO Photometric all-sky survey (APASS) DR6
\citep{Henden14}, as reported in the UCAC4 Catalogue \citep{Zacharias12}.
We also record the $J$, $H$, and $K$ magnitudes and their uncertainties
as found in the 2MASS All-Sky Catalog of Point Sources \citep{Cutri03}
and the $W1-W3$ WISE magnitudes and uncertainties from the ALLWise Data
Release \citep{Cutri13}.
For all except two of our targets, the $W4$ band is only an upper limit,
and in the remaining two cases, the photometric uncertainity in $W4$ is at least an
order of magnitude larger than those in $W1-W3$, so we neglect $W4$.
These data are reported in Table 1, and a color-color diagram showing the
$r-J, J-K$ colors of our candidates is included as Fig. \ref{fig:photometry}.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.5\textwidth]{f1}}
\caption{Color-color diagram displaying $r-J, J-K$ photometry for targets
observed by \kep\ during the original
mission (black), with our \KT\ \Ci\ planet candidates overlaid (fuchsia).
Also included is the location of the Sun (yellow).
Nearly all of our candidates have photometry consistent with later spectral
types than G2.}
\label{fig:photometry}
\end{figure}

\subsection{Stellar Models}
\label{sec:stellarparams}
To convert the observed photometric data into physical properties for each
star, we used the publicly available \texttt{isochrones} Python module\footnote{
\url{http://github.com/timothydmorton/isochrones}}, a general-purpose
interpolation tool for stellar model grids.
This software does trilinear interpolation in mass--age--\feh\ space for any
given set of model grids, thus being able to predict the value for any
physical or photometric property provided by the models at any values of
mass, age, and \feh\ within the boundaries of the grid.

This enables a set of observed properties ($\left\{ x_i, \sigma_i \right\}$),
either spectroscopic, photometric,
or both, to define a likelihood function that may be sampled in
an MCMC chain:
\begin{equation}
\label{eq:ischochronelhood}
\ln \mathcal L(\boldsymbol{\theta}) = -\frac{1}{2} \displaystyle \sum_i \frac{\left(x_i -
  I_i\left(\boldsymbol{\theta} \right)\right)^2}{\sigma_i^2},
\end{equation}
where $I_i(\boldsymbol{\theta})$ is the isochrone model prediction of property
$i$ at the given parameters $\boldsymbol{\theta}$.
If the observed properties include any apparent magnitudes, then
$\boldsymbol{\theta}$ includes distance and $A_V$ extinction in addition
to mass, age, and \feh.

In this work, we use grids from the Dartmouth Stellar
Evolution Database \citep{Dotter08} at solar values of \afe=$0.0$ and
helium abundance $Y=0.2741$, which come packaged with the \texttt{isochrones}
module.
We then run an MCMC chain using \texttt{emcee} \citep{Foreman-Mackey12},
an implementation of the affine-invariant ensemble sampler of
\citet{Goodman10}
for each host star conditioned on the observed
photometric properties as presented in Table 1.
We include a prior on stellar metallicity representative of the observed metallicities
of stars within 1 kpc of the Sun, following the results of \citet{Hayden15}.

Such a scheme enables us to infer the statistical uncertainties on the
mass, radius, and effective temperature.
However, we are subject to biases induced by systematics in the models themselves.
There is some evidence that the Dartmouth models may under-predict
radii by $\sim 15\%$ when compared to other methods \citep{Newton15,
Montet15}.
Such an effect may be the result of the Dartmouth model reliance on BT-Settl
atmospheres, which have been shown to predict near-IR colors that are too blue
\citep{Thompson14}.

In order to try to encapsulate the effect of model uncertainties, we also run
MCMC chains with photometric errors inflated artificially by a factor of three.
We provide full chains using both inflated and un-inflated uncertainties,
though in the rest of this work (and in Table 2) we adopt
the results obtained using the inflated values.

\subsection{SpeX Spectroscopy}
\label{Spexobs}

A near-infrared spectrum of EPIC 201912552 was obtained using the updated SpeX
(uSpeX) spectrograph \citep{Rayner03} on the NASA Infrared Telescope Facility
(IRTF) on January 29 2015 (UT).
SpeX observations were taking using the short cross-dispersed mode and the
0.3$\times15\arcsec$ slit, which provides simultaneous coverage from 0.7
to 2.5$\mu$m at $R\simeq2000$.
The target was observed at two positions along the slit to subsequently subtract
the sky background. Eight spectra were taking following this pattern, which provided
a final S/N of $>150$ per resolving element.
The spectrum was flat fielded, extracted, wavelength calibrated, and stacked
using the \textit{SpeXTool} package \citep{Cushing04}.
An A0V-type star was observed immediately after
the target, which was used to create a telluric correction using the
\textit{xtellcor} package \citep{Vacca03}.

An optical spectrum was obtained using the SuperNova Integral Field Spectrograph
\citep[SNIFS,][]{Aldering02,Lantz04} on the University of Hawai'i
2.2m telescope on the night of January 30 2015.
SNIFS provides simultaneous coverage from 3200\AA--9700\,AA\ at a resolution
of $\simeq1000$. Final S/N of the spectrum was $>100$ pre resolving element
in the red ($\sim6000$\AA).
Details of the SNIFS reduction, including dark, bias, and flat-field corrections,
cleaning the data of bad pixels and cosmic rays, and extraction of the
one-dimensional spectrum are described in \citet{Bacon01} and
\citet{Aldering06}.
Flux calibration was performed using a separate pipeline described in \citet{Mann15}.

\teff\ was calculated by comparing our optical spectra with the CFIST
suite\footnote{http://phoenix.ens-lyon.fr/Grids/BT-Settl/CIFIST2011/} of the BT-SETTL
version of the PHOENIX atmosphere models \citep{Allard13}, which gave a temperature
of 3503 $\pm$ 60\,K.
More details of this procedure are given in \citet{Mann14} and
\citet{Gaidos14}.
This method was used because it is known to accurately reproduce empirical
\teff\ values from long-baseline optical interferometry \citet{Boyajian12}.

Metallcity was determined using the procedures from \citet{Mann13a}, in which the
authors provide empirical relations between atomic features and M dwarf
metallicity, calibrated using wide binaries.
We adopted the weighted mean of the $H-$ and $K-$band calibrations,
which yielded a metallicity of 0.09$\pm$0.09.

We combined the derived \teff\ and [Fe/H] values with the empirical
\teff-[Fe/H]-$R_*$ relation from \citet{Mann15} to compute a radius.
Accounting for measurement and calibration errors in [Fe/H] and \teff\ we calculated
a radius 0.394$\pm0.038R_\odot$.
We use these parameters instead of the derived photometric properties for this target.

The full list of stellar parameters adopted in this paper is included in
Table 2.
% For some reason I can't get it to identify the table \ref here.
% If anyone can figure it out, be my guest!

\section{Planet Properties}

In \paperit, we present only parameters directly observable from the \KT\
light curve itself: the period, time of transit center, and transit depth.
With stellar properties now in hand, we can convert these observational
results into fundamental parameters of each planet candidate.

For each candidate, we fit a transit and systematic model simultaneously
to the data.
The transit model is an exposure-time integrated, two-parameter limb darkened 
light curve \citep{Mandel02,Kipping10b}.
We parameterize... \dfmtodo{What are the parameters in the model?}

This method enable us to marginalize over the parameters of the
data-driven model of the instrumental systematics described in \paperit.

The results are given in Table 3.

\dfmtodo{Add more to this section.}

In this analysis, we assume the dilution caused by additional stars contributing
flux into the aperture is negligible for nearly all systems.
Given the location of the \Ci\ field at a high galactic latitude, we expect low
contamination by background giants. 
Nevertheless, this assumption may not be valid for all systems.
Any contamination unaccounted for, as may happen if any of these stars are
actually unresolved binaries, would cause us to underestimate the radii
of any planets we detect.
Therefore, high-contrast adaptive optics imaging of any systems should be
obtained before any of these planets are placed on a planet mass-radius
diagram.

One system, EPIC 201828749, is known to be a binary through AO imaging
as discussed in \ref{sec:AO}. \btmtodo{More text...}

\section{False Positive Analysis}

As has been well known from the early days of transit searches, it is
possible for an eclipsing binary star to masquerade as a transiting
planet if (a) it is a highly grazing eclipse, or (b) the binary system
comprises only a small fraction of the total light in the photometric
aperture, resulting in a diluted eclipse depth.
When possible, such astrophsycial false positive scenarios are traditionally
ruled out by detailed follow-up observations, often a combination of
high-resolution imaging and radial-velocity measurements.
However, the \kep\ mission, with its thousands of planet candidates
around mostly faint stars, necessitated a paradigm shift---a move
toward probabilistic interpretation of transit signals, rather than
comprehensive follow-up of each individual candidate.

\citet{Morton12} presented an automated method to calculate the
probability that a planet candidate might be caused by an
astrophysical false positive.
This method uses population simulations of the light curves of
of false positive scenarios and true transiting planets to quantify
the typical shape of each, comparing to the observed transit shape.
It then combines this information with observationally motivated prior
assumptions about the populations of field stars, the properties of
multiple star systems, and the occurrence rate of planets as determined
from \kep\ \citep{Fressin13}, in order to determine the probability that
the observed signal may be a false positive.
Similar in spirit to other published methods of
probabilistic validation, such as Blender \citep{} and PASTIS
\citep{}, it has the advantage of being computationally less demanding
and fully automated, and thus easily applied in batch to a large
number of candidates.

In this work, we use
\texttt{vespa}\footnote{\url{http://github.com/timothydmorton/vespa}},
a new general-purpose implementation of the \citet{Morton12} procedure,
to calculate false positive probabilities (FPPs) for each of these
\KT\ candidates.  We impose the following constraints on false positive
scenarios:
\begin{itemize}
\item A chance-aligned eclipsing binary system may reside anywhere within
an 8 arcsecond radius from the target star.
\item The maximum allowed depth of a potential secondary eclipse event
is the most significantly detected signal at the same
period of the planet candidate, once the primary transit is masked out.
\item Blended stars must be allowed by the available adaptive optics
and archival imaging data (discussed in detail below).
\end{itemize}

Table 4 summarizes the results of these
calculations, presenting the relative probability for each candidate
to be caused by any of three false positive scenarios: an undiluted
eclipsing binary (EB), a hierarchical triple eclipsing binary (HEB),
and a chance-aligned background(/foreground) eclipsing binary (BEB).

We find that \Nfp\ of the presented candidates FPP $>$90\%;
these we consider to be likely false positives.
On the other hand, we find that \Nvalidated\ candidates
have FPP $<$ 1\%; these we consider to be validated planets.
The remaining \Nlimbo\ candidates have FPPs neither high enough to discard
nor low enough to securely validate.

We emphasize that the majority of these planets were validated based
solely on the transit signal, with no need for follow-up imaging.  This
demonstrates the utility of \texttt{vespa} analysis, which will be
crucial to interpreting future \KT\ detected candidates and prioritizing
follow-up observing efforts.

\subsection{Adaptive Optics Imaging}
\label{sec:AO}

This is Brendan.

Images for all systems AO data was obtained for is shown in Fig. \ref{Fig:AO}.

\begin{figure*}[htbp]
\centerline{\includegraphics[width=1.0\textwidth]{f2}}
\caption{AO data for all systems imaged.
The main frame for each single system shows the deep, saturated image.
The inset for each single system shows a shallower, unsaturated image to better
identify companions at close projected orbital separations.
For the two systems with imaged companions, EPIC 201546283 and 201828749,
only unsaturated frames are collected.
All subplots, including insets, are plotted on the same scale.}
\label{Fig:AO}
\end{figure*}

\anytodo{201828749 has an AO companion, flux ratio 6:1 or so. We call it a planet.
It's a very flat bottom thing, with fairly short ingress and egress: the transit shape
is way more planet than EB so I'm comfortable with this designation. We need
to talk about this though.}

\subsection{Known Background Stars}
The PHARO AO system has a field of view of NUMBER arcseconds.
\btmtodo{Check with Brendan to find this number}.
Each \KT\ pixel is a square of $3.98 \times 3.98$ arcseconds.
A background eclipsing binary within a few \KT\ pixels of our target
stars could mimic a transit signal inside our aperture while evading detection
by PHARO.
Such wide eclipsing binaries should appear in seeing-limited ground-based
surveys.

To investigate the possibility that such wide companions exist,
we query the ninth data release of the Sloan Digital Sky Survey
\citep[SDSS DR9,][]{Ahn12}.
For each target, we return
all stars within 12\arcsec with $\Delta r < 8$ relative to each object
of interest.
Of the 31 stars in our sample, four have such a companion.
Unlike the original \kep\ field, the field for \KT\ \Ci\ is
well out of the galactic plane, so the rate of giant, distanct background
stars is significantly lower.
We include a list of these potential contaminants as Table 5.

In Table 5, the ``maximum depth'' column represents
the maximum observed ``transit'' depth if the transit were actually caused
by a total eclipse of the hypothetical background binary system, inducing
a 50\% flux decrement in the background star's apparent brightness.

For EPIC 201465501 and 201779067, the observed transit depth is
larger than each of these values, so we reject the possibility that
these events could be the eclipses of the background binaries detected
here.

For EPIC 201546283, the observed transit depth is nearly identical to
the maximum false positive depth.
The photometry for this potential contaminant is consistent with that
of an M3V star; for this object to cause the observed eclipse it must
be a total eclipse of two M3V stars.
However, in that case, the observed event would be exactly V-shaped,
with second and third contact occurring simultaneously.
Since the transit is not V-shaped, it must be the result of
a relatively small body passing in front of a relatively large body.
Therefore, it cannot be caused by this observed background dwarf star.

For EPIC 201295312, the transit depth is significantly shallower than
the maximum false positive depth, so we cannot immediately
reject the hypothesis that the transit is actually a false positive
induced by a background eclipsing binary.
However, from the transit shape alone we can place a FPP of this
system of $1.1 \times 10^{-4}$.

SDSS is 95\% complete at $r=22.2$ and the telescope has a point
spread function of $1\farcs4$. In all cases except the four above, we
treat the SDSS data as providing a contrast curve at wide separations
down to a limiting magnitude of $r=22.2$.

\subsection{Archival Imaging}

For the stars with AO nondetections, there is still the
possibility that a background binary could be positioned
directly behind the target star, evading detection.
The probability is small, given the $0\farcs1$ diffraction limit
of P200 at $2\mu$m, but nonzero.
We can rule out the possibility of such chance alignments with
archival imaging data.

For all targets with AO, we can confirm the planetary nature of the signal
without this extra step, but we include it for EPIC 201912552 for
completeness, considering thee additional follow-up this target may warrant.
We download data from the POSS I and II surveys, which imaged
EPIC 201912552 in 1955 and 1989, respectively.
This star has a proper motion of $-74 \pm 2$ mas yr$^{-1}$ in RA and
$-122 \pm 4$ mas yr$^{-1}$ in declination, so it has moved by $8.35\arcsec$
since the POSS I data were collected.
As shown in Figure \ref{fig:archival}, we do not detect any targets at the
present-day location of EPIC 201912552 in either of the POSS surveys.

\begin{figure*}[htbp]
\centerline{\includegraphics[width=1.0\textwidth, trim=0 6.30cm 0 6.30cm, clip=true]{f3}}
\caption{Imaging for EPIC 201912552 from 1955 to 2002. There is no background
object directly behind the present day location of the target (red circle) that could be missed
by the AO observations, while modern SDSS imaging can rule out wide companions
that may have been missed at wide separations, beyond the AO field of view.
All subplots are on the same scale.}
\label{fig:archival}
\end{figure*}

For this target, we can extend our contrast curves to zero
present-day orbital separation and rule out the possibility that
these transit events are caused by a background eclipsing binary.
By combining present-day seeing-limited photometric survey data,
adaptive optics imaging, and archival photometry, the only stellar companions
we would not detect would be those that are gravitationally bound to the target
star and positioned in their orbits so that their projected separation is
smaller than the diffraction limit of the P200 telescope.
Such an alignment would require the orbital inclination of the binary to be
nearly $90^\circ$ and the phase $\varpi + \theta \approx \pi/2$ or $3\pi/2$.
While we cannot fully rule out this possibility, such an alignment is unlikely
to occur, and highly unlikely to cause enough dilution to call the planetary
nature of this signal into question.

\subsection{TRES Radial Velocities}

We observed EPIC 201912552 on (all UT) February 04 2015 and February 24 2015
with the Tillinghast Reflector Echelle
Spectrograph (TRES) on the 1.5 m Tillinghast Reflector at the Fred L.
Whipple Observatory.
These dates were chosen to be near the times of largest RV variations,
corresponding to phases of 0.72 and 0.32 relative to the time
of transit.
The spectra were taken with a resolving power of $R=44,000$ and
integration times ranging from 2800 to 3600 seconds, resulting in
signal-to-noise ratios between 17 and 29 per resolution element.

The spectra were extracted as described in \citet{Buchhave10}.
The relative RVs were derived by cross-correlating the spectra against the
strongest observed spectrum (in this case, the first) over the wavelength
range 4700 - 6800 Angstroms.
We selected 19 echelle orders in the analysis, being careful to reject
orders with telluric absorption lines, fringing in the far red and those
with very low SNR in the blue.

The two observed spectra have RVs that differ by $47.1 \pm 41.9$ m s$^{-1}$.
A deviation at this magnitude can only be explained by the presence of
a planetary system and instrumental noise, eliminating the possibility the
transiting planet is actually a false positive.

\section{Potentially Interesting Systems}
\subsection{A Mini-Neptune with Earthlike Insolation}

Of particular interest is EPIC 201912552.

By combining archival and modern seeing-limited data with adaptive optics
imaging, we can exclude the possibility these transit events are caused by
a background eclipsing binary.
The apparent transits must be caused by an object co-moving with EPIC 201912552;
radial velocities eliminate the possibility the companion is nonplanetary.
Therefore, we confirm the planetary nature of this system.

This star is an M2.8 dwarf at a distance of $34\pm4$ pc.
Of our planet candidate hosts, only EPIC 201367065 \citep[originally
discovered by][]{Crossfield15} is brighter in $K$-band.
This star is only 0.1 magnitudes fainter in $K$ than GJ\,1214
\citep{Charbonneau09}.
Due to the relative brightness of the host star, this target is likely
to become a prime target for atmospheric characterization studies
and is ideal as a target for future space-based missions such as JWST.

Unlike GJ\,1214b, EPIC 201912552b is not highly irradiated.
Instead, it is at an reduced semimajor axis \ars$ = 83.2 \pm 9.1$.
Its equilibrium temperature is then, assuming zero albedo, $T_{eq} = 273 \pm 15$
K, meaning its bulk insolation is $131 \pm 30$ percent that of the Earth's.
Although the planet is likely too large to be rocky \citep{Rogers14},
its atmosphere is likely to be the focus of many future observations, providing
a cool analogue to the highly irradiated planets of a similar size found by
\kep.

\subsection{Two Candidates, One Ephemeris}

EPIC 201569483 was originally listed in \paperit\ as a planet candidate.
The transit events have a depth of approximately 7\%.
In this work, we conclude this system is more likely than not a false
positive, with an FPP of 0.560.
The most likely scenario is that of an eclipsing binary, which would
not be surprising given the transit depth.

It was later realized that EPIC 201555883 has a period, time of transit,
and transit duration consistent with EPIC 201569483.
Such effects are not uncommon in \kep\ data.
\citet{Batalha13} found $\sim 25$ planet candidates to be false positives based on matching
ephemerides between candidates.
In that paper, the authors limited their search to candidates within $20\arcsec$ of other
objects of interest.
\citet{Coughlin14} expanded on this method, identifying 685 KOIs as false positives and
outlining four physical reasons why these anomalies may occur.
As those authors explain, period-matching candidates can be the result of direct PRF
contamination, antipodal reflection across the Schmidt telescope, CCD crosstalk
across a \kep\ module, or ``column anomalies'' across one column of the CCD.

EPIC 201555883 is a unique case in that it does not appear to fall under any of these
cases. It falls on module 23, while EPIC 201569483 is on module 8, neither 180 degrees
away from or on the same column as this candidate.
Moreover, there is not any evidence of a mechanism that could cause a third star to
induce both the appearance of a $7\%$ eclipse on one module and an additional
transit event on a different module.

Formally, our false positive validation scheme gives a FPP of $9.5 \times 10^{-4}$,
reaching our threshold for planet validation.
In this case, given the period matching between the two events, we refrain from
declaring this system a planet.
However, since we cannot explain how such a false positive would occur as a result
of an instrumental systematic in the \kep\ telescope, we also refrain from declaring
this target a definite false positive.
Additional follow-up observations of both targets are clearly warranted to better
understand a possible systematic that has not previously been observed in
either \kep\ or \KT.

\subsection{Multiple Planet Systems}
Five of the systems identified by \paperit\ are candidates to be multiply-transiting
systems.
One of these is EPIC 201367065, a three-planet system originally announced by
\citet{Crossfield15}.
Another of these is EPIC 201505350 \citep{Armstrong15b}, a two-planet system with the
orbital periods of the two planets near a 3:2 period commensurability.
The remaining three are all representative of the multiple-planet systems observed by
\kep \citep{Lissauer11b, Fabrycky14}.
Two of the systems are near a period commensurability and all three consist of
mini-Neptune sized planets.
Like in the \kep\ field, the false positive rate we measure for the multiple planet sample
here ($0/10$) is lower than for the single systems \citep{Lissauer12, Rowe14}.

We do not detect any significant transit timing variations (TTVs) in any of these systems from
the \KT\ data alone.
EPIC 201338508 would be expected to have a TTV period of 117 days, but
is likely too far from commensurability to have an observable TTV signal.
EPIC 201445392 is expected to have a TTV period of 234 days, so this system may be a candidate
for additional follow-up to constrain the system masses dynamically.
The transiting planets orbiting EPIC 201754305 are near a 5:2 period commensurability.
There is no evidence from \kep\ of an abundance of planets near this period ratio, and so this
may be coincidence.
Follow-up observations may be warranted to search for an additional planet in this system
forming a resonant chain, similar to those observed around other stars \citep[e.g.][]{Swift13,
Campante15}.

\subsection{Systems Orbiting Bright Stars}

One of the primary goals of \KT\ is the detection of transiting planets around bright stars that
can be followed up from the ground or with future space-based observatories such as JWST
\citep{Howell14}.
Of our sample, two systems orbit stars with $K < 9$: EPIC 201367065 \citep{Crossfield15} and
EPIC 201912552.
An additional four targets orbit stars with $K <10$.
These targets are ideal for ground-based followup, and may be useful targets for Spitzer and
JWST to probe planetary atmospheres.

\section{Results and Discussion}

We have presented stellar parameters for all planet candidates systems observed by
\paperit, validating \Nvalidated\ of the 36 candidates as bona fide planets and
\Nfp\ as bona fide false positives.
Of these planets, 4 have been previously validated in other works, while 17 are
validated here for the first time.

With the exception of one object, all of the stellar parameters are derived from comparing
photometric observations to the Dartmouth stellar evolution models.
As a result, both the stellar and planet parameters are subject to systematic biases induced
by discrepancies between the models and reality.
Even more dangerously, any region of parameter space not probed by the models would
be entirely missed in our stellar modeling.

As stated in \textsection\ref{sec:stellarparams}, we infer stellar parameters using the
Dartmouth models, with a mass upper limit of $3.7 M_\odot$.
Given the location of the \Ci\ field with respect to the galactic plane, we expect a low
contamination rate by halo giants.
However, there are a few stars that are possibly miscategorized giants.
EPIC 201257461 is identified in the RAdial Velocity Experiment (RAVE) Fourth Data Release
as a target with $\logg = 3.18$, consistent with that of a giant star.
It is therefore perhaps not surprising that this system is flagged as a false positive based on
the transit shape.
EPIC 201929294, with a proper motion of $0.010\arcsec$ yr$^{-1}$ \citep{Zacharias12}  has
a reduced proper motion consistent with that of a giant star.
While suggestive, this observation does not definitively provide information about the
nature of this target.
Likewise, our false positive analysis is neither able to confirm nor exclude the planetary
nature of the observed transit signal around this target.

The planets we confirm in this paper, like the planets found in the original \kep\ mission,
span a wide range of parameter space.
They are at distances ranging from $34$ to $500$ pc, have radii ranging from $1.3$ to
$5.3 R_\odot$, and orbital periods ranging from 5.0 to 50.3 days.
Like the original mission, we find significantly more small planets than large planets.

Unlike the original mission, however, we find that nearly all of our confirmed planets
are around stars less massive than the Sun.
This difference is a result of both the \Ci\ field and the target selection process.
\Ci\ is at a significantly higher galactic latitude than the original \kep\ mission,
meaning there is a much lower number density of targets at large distances.
A magnitude limited survey pointed in this direction would then have much in
common with a volume limited survey near the galactic plane, as massive stars
at kiloparsec distances are relatively unlikely likely to exist.
Therefore, the pool of targets that could be selected for \Ci\ contains a larger
fraction of subsolar stars.

Low-mass stars, particularly M dwarfs, are also a specific focus of the \KT\ mission.
One of the primary goals of the Kepler mission was to ``Determine the abundance of
terrestrial and larger planets in or near the habitable zone of a wide variety of spectral
types of stars'' \citep{Batalha13} However, $\sim70\%$ of Kepler's target stars had
masses within 20\%
of the Sun's, while 70\% of the stars in the Galaxy have less than 50\% the mass of the
Sun \citep{Brown11}.
\KT\ will fulfill the promise of \kep, with the goal of providing a yield of small planets
around bright, small stars to facilitate follow-up measurements \citep{Howell14}.
This is clear from the \KT\ target selection process, with thousands of K and M dwarfs
being selected in each campaign.
Based on these plans, we expect that \KT\ will detect hundreds of planets during its
lifetime, with the majority being mini-Neptunes and super-Earths around stars
less massive than the Sun.

\anytodo{Does anyone feel we should include anything else in the discussion?}

\anytodo{We should have a figure of many subplots, of phase-folded
transits for all the confirmed planets.
To ease Dan's mind, we will say that they are the product of
some fidicual transit model and are only presented here for illustrative
purposes.}

\acknowledgements
We thank Eric Agol (UW), Dan Huber (Sydney) and Jeff Coughlin (SETI) for
conversations and suggestions which improved the quality of this manuscript.

We are grateful to the entire Kepler team, past and present.
Their tireless efforts were all essential to the tremendous success of the mission and the
successes of \KT\, present and future.

Some of the data presented in this paper were obtained from the Mikulski
Archive for Space Telescopes (MAST).
STScI is operated by the Association of Universities for Research
in Astronomy, Inc., under NASA contract NAS5--26555.
for MAST for non--HST data is provided by the NASA Office of Space
Science via grant NNX13AC07G and by other grants and contracts.

This paper includes data collected by the \kep\ mission.
Funding for the \kep\ mission is provided by the NASA Science
Mission directorate.

This paper includes data collected by the Sloan Digital Sky Survey.
Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation,
the Participating Institutions, the National Science Foundation, and the
U.S. Department of Energy Office of Science. The SDSS-III web site is
http://www.sdss3.org/.
SDSS-III is managed by the Astrophysical Research Consortium for the
Participating Institutions of the SDSS-III Collaboration including the
University of Arizona, the Brazilian Participation Group, Brookhaven
National Laboratory, Carnegie Mellon University, University of Florida,
the French Participation Group, the German Participation Group, Harvard
University, the Instituto de Astrofisica de Canarias, the Michigan
State/Notre Dame/JINA Participation Group, Johns Hopkins University,
Lawrence Berkeley National Laboratory, Max Planck Institute for
Astrophysics, Max Planck Institute for Extraterrestrial Physics, New
Mexico State University, New York University, Ohio State University,
Pennsylvania State University, University of Portsmouth, Princeton
University, the Spanish Participation Group, University of Tokyo,
University of Utah, Vanderbilt University, University of Virginia,
University of Washington, and Yale University.

B.T.M. is supported by the National Science Foundation Graduate Research
Fellowship under Grant No. DGE-€1144469.
J.A.J. is supported by generous grants from the David and Lucile Packard
Foundation and the Alfred P. Sloan Foundation.

D.F.M. and D.W.H. were partially supported by the National Science Foundation
(grant IIS-1124794), the National Aeronautics and Space Administration (grant
NNX12AI50G), and the Moore-Sloan Data Science Environment at NYU.

T.D.M.~is supported by the National Aeronautics and Space Administration (grant
NNX14AE11G).

{\it Facilities:} \facility{Kepler},  \facility{PO:Hale (PHARO)}, \facility{FLWO:1.5m},
\facility{IRTF:SpeX}

\bibliography{exopapers}

% \begin{turnpage}
\include{table_photometry}
% \end{turnpage}
\include{table_stellarprops}
\include{table_planets}
\include{table_fpp}

\include{table_bg}

\end{document}