### 2015-03-26 table_fpp.tex
fpp table edits.  latex formatting error, sorry..

\clearpage
%\LongTables
\begin{deluxetable*}{ccccccccc}
\tablewidth{0pt}
\tabletypesize{\scriptsize}
\tablecaption{False Positive Probability Calculation Results}
\label{Table:FPP}
\tablehead{
\colhead{Candidate} &
\colhead{$\delta_{\rm sec, max}$ [ppt]\tablenotemark{1}} &
\colhead{AO?\tablenotemark{2}} &
\colhead{${\rm Pr}_{\rm EB}$} &
\colhead{${\rm Pr}_{\rm BEB}$} &
\colhead{${\rm Pr}_{\rm HEB}$} &
\colhead{$f_{p}$\tablenotemark{3}} &
\colhead{FPP} &
\colhead{Disposition}
}
\startdata
201208431.01 & $0.51$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $7.0\times10^{-4}$ & $0.21$ & $7.0\times10^{-4}$ & Planet \\
 \color{red} 201257461.01  & \color{red}  $0.59$  & \color{red}   -  & \color{red}  $8.8\times10^{-3}$  & \color{red}  $0.940$  & \color{red}  $0.051$  & \color{red}  $0.17$  & \color{red}  $1.000$  & \color{red}  FP\\
201295312.01 & $0.04$ &  Y & $< 10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.21$ & $< 10^{-4}$ & Planet \\
201338508.01 & $0.63$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $3.2\times10^{-3}$ & $0.21$ & $3.2\times10^{-3}$ & Planet \\
201338508.02 & $0.33$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $1.8\times10^{-4}$ & $0.22$ & $1.8\times10^{-4}$ & Planet \\
201367065.01 & $0.15$ &  - & $0.717$ & $< 10^{-4}$ & $0.044$ & $0.07$ & $0.761$ & Planet\tablenotemark{c} \\
201367065.02 & $0.67$ &  - & $3.2\times10^{-3}$ & $< 10^{-4}$ & $4.2\times10^{-3}$ & $0.17$ & $7.4\times10^{-3}$ & Planet\tablenotemark{c} \\
201384232.01 & $0.44$ &  - & $1.2\times10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.16$ & $1.4\times10^{-4}$ & Planet \\
201393098.01 & $0.52$ &  - & $1.6\times10^{-3}$ & $< 10^{-4}$ & $5.5\times10^{-4}$ & $0.19$ & $2.2\times10^{-3}$ & Planet \\
201403446.01 & $0.18$ &  Y & $< 10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.22$ & $< 10^{-4}$ & Planet \\
201445392.01 & $0.26$ &  - & $1.3\times10^{-3}$ & $< 10^{-4}$ & $0.012$ & $0.19$ & $0.013$ & Candidate \\
201445392.02 & $0.18$ &  - & $0.023$ & $< 10^{-4}$ & $0.254$ & $0.21$ & $0.277$ & Candidate \\
201465501.01 & $0.68$ &  - & $0.491$ & $0.022$ & $9.1\times10^{-3}$ & $0.20$ & $0.522$ & Candidate \\
201505350.01 & $2.69$ &  - & $0.583$ & $< 10^{-4}$ & $0.024$ & $0.07$ & $0.607$ & Planet\tablenotemark{d} \\
201505350.02 & $0.70$ &  - & $0.034$ & $< 10^{-4}$ & $2.3\times10^{-4}$ & $0.07$ & $0.034$ & Planet\tablenotemark{d} \\
201546283.01 & $0.15$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.06$ & $1.1\times10^{-4}$ & Candidate\tablenotemark{a} \\
201549860.01 & $0.18$ &  - & $7.5\times10^{-3}$ & $0.014$ & $0.011$ & $0.22$ & $0.032$ & Candidate \\
 \color{red} 201555883.01  & \color{red}  $0.94$  & \color{red}   -  & \color{red}  $< 10^{-4}$  & \color{red}  $< 10^{-4}$  & \color{red}  $3.6\times10^{-3}$  & \color{red}  $0.10$  & \color{red}  $3.6\times10^{-3}$  & \color{red}  FP\tablenotemark{b}\\
201565013.01 & $1.69$ &  - & $0.460$ & $0.391$ & $0.035$ & $0.06$ & $0.887$ & Candidate \\
 \color{red} 201569483.01  & \color{red}  $2.06$  & \color{red}   -  & \color{red}  $0.437$  & \color{red}  $0.562$  & \color{red}  $< 10^{-4}$  & \color{red}  $0.00$  & \color{red}  $0.999$  & \color{red}  FP\\
201577035.01 & $0.14$ &  Y & $3.0\times10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.07$ & $3.6\times10^{-4}$ & Planet \\
201596316.01 & $0.45$ &  - & $0.017$ & $< 10^{-4}$ & $3.4\times10^{-4}$ & $0.20$ & $0.017$ & Candidate \\
201613023.01 & $0.08$ &  Y & $9.8\times10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.20$ & $9.8\times10^{-4}$ & Planet \\
201617985.01 & $0.27$ &  - & $0.166$ & $< 10^{-4}$ & $0.053$ & $0.19$ & $0.219$ & Candidate \\
201629650.01 & $0.43$ &  - & $1.5\times10^{-3}$ & $< 10^{-4}$ & $2.6\times10^{-4}$ & $0.18$ & $1.8\times10^{-3}$ & Planet \\
201635569.01 & $0.79$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $0.043$ & $0.06$ & $0.043$ & Candidate \\
 \color{red} 201649426.01  & \color{red}  $3.10$  & \color{red}   -  & \color{red}  $0.856$  & \color{red}  $0.142$  & \color{red}  $1.6\times10^{-4}$  & \color{red}  $0.00$  & \color{red}  $0.998$  & \color{red}  FP\\
201702477.01 & $0.70$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $4.0\times10^{-3}$ & $0.06$ & $4.0\times10^{-3}$ & Planet \\
201736247.01 & $0.42$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $4.6\times10^{-4}$ & $0.18$ & $4.9\times10^{-4}$ & Planet \\
201754305.01 & $0.65$ &  - & $2.5\times10^{-3}$ & $< 10^{-4}$ & $3.2\times10^{-3}$ & $0.21$ & $5.7\times10^{-3}$ & Planet \\
201754305.02 & $0.38$ &  - & $1.9\times10^{-3}$ & $< 10^{-4}$ & $1.7\times10^{-3}$ & $0.22$ & $3.6\times10^{-3}$ & Planet \\
 \color{red} 201779067.01  & \color{red}  $1.97$  & \color{red}   -  & \color{red}  $0.935$  & \color{red}  $0.063$  & \color{red}  $3.2\times10^{-4}$  & \color{red}  $0.00$  & \color{red}  $0.998$  & \color{red}  FP\\
201828749.01 & $0.39$ &  Y & $2.6\times10^{-3}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.17$ & $2.6\times10^{-3}$ & Candidate\tablenotemark{a} \\
201855371.01 & $0.62$ &  - & $< 10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.22$ & $< 10^{-4}$ & Planet \\
201912552.01 & $0.47$ &  Y & $< 10^{-4}$ & $< 10^{-4}$ & $< 10^{-4}$ & $0.21$ & $< 10^{-4}$ & Planet \\
 \color{red} 201929294.01  & \color{red}  $3.12$  & \color{red}   -  & \color{red}  $< 10^{-4}$  & \color{red}  $< 10^{-4}$  & \color{red}  $2.8\times10^{-4}$  & \color{red}  $0.07$  & \color{red}  $2.9\times10^{-4}$  & \color{red}  FP\tablenotemark{b} 
 &
\enddata
\tablecomments{Results of the \texttt{vespa} astrophysical 
false positive probability calculations for all candidates.  
Likely false positives (FPP $> 0.9$, or otherwise designated)
 are marked in red.  
Candidates are declared to be validated planets if FPP $< 0.01$.  
EB, BEB, and HEB refer to the three considered astrophysical 
false positive scenarios, and the relative probability of 
each is listed in the appropriate column.  Planets previously 
identified in the literature are marked as planets even if FPP is not 
below 1\% threshold.}
\tablenotetext{1}{Maximum depth of potential secondary eclipse signal.}
\tablenotetext{2}{Whether adaptive optics observation is presented in this paper.}
\tablenotetext{3}{Integrated planet occurrence rate assumed between 0.7$\times$ and 1.3$\times$ the candidate radius}
\tablenotetext{a}{Despite low FPP, returned to candidate status 
out of abundance of caution due
to secondary star detection in or near photometric aperture.  Most of these we do anticipate to be validated once the specific false positive scenarios are considered.}
\tablenotetext{b}{Declared a false positive due to noise modeling systematics (see \S\ref{sec:systematics}).}
\tablenotetext{c}{Identified as planets by \citet{Crossfield15}.}
\tablenotetext{d}{Identified as planets by \citet{Armstrong15b}.}
\end{deluxetable*}